Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-12 paper/article.tex
Paper progress

TODO: table of raw healthy vs lesioned vs recovered vs pct recovered

@@

\subsection{Experiment 1: AIP 75\% lesion}
This experiment measures the effect of a loss of task representation neural
machinery on the ability of the co-processor to succeed. In the non-coadaptive
version of this experiment, the loss of neurons involved in encoding the
visual inputs results in the simulated brain being unable to condition its
grasp for the object shape. Likewise, the co-processor doesn't have sufficient
information to improve performance. The lesioned circuit largely

In the co-adaptive case,

See Fig. \ref{fig:results}.

\subsection{Experiment 2: M1 75\% lesion}
\subsection{Experiment 3: F5 and M1 connectivity lesion}
\subsection{Experiment 4: M1 lesion with pre-recovery}

TODO: summary of conclusions that will go here:

TODO:
* AIP + coadaptation shows that we can adapt to changes in the input mappings.
* M1 + coadaptation shows we can adapt to changes in the output mappings
* Connection lesion shows we can act as a bridge

\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery with and without co-adpation. The co-processor is having trouble finding a better solution than this, due to loss in brain functionality. Increasing the rate of brain co-adaptation may show a better solution.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Can co-adapt with the brain, as it changes
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. Initial result: the co-adaptive case found a better solution. The co-processor couldn't improve yet on the pre-recovery case. The solution it found was (basically) to ignore the inputs.
\end{itemize}

>>>>

TODO: label lines/colors

\begin{table}[]
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Experiment}  & \textbf{\begin{tabular}[c]{@{}l@{}}Lesioned\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task\\ loss\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Min task val\\ loss\end{tabular}} & \textbf{Pct recov} & \textbf{\begin{tabular}[c]{@{}l@{}}Pct recov\\ val\end{tabular}} \\ \hline
AIP No-coadapt & 0.004507                                                          & 0.004599                                                         & 0.004658                                                             & -2.32\%            & -3.82\%                                                          \\ \hline
AIP Coadapt    & 0.004507                                                          & 0.001575                                                         & 0.001502                                                             & 74.41\%            & 76.26\%                                                          \\ \hline
M1 No-coadapt  & 0.021136                                                          &                                                                  &                                                                      & 102.75\%           & 102.75\%                                                         \\ \hline
M1 Coadapt     & 0.021136                                                          & 0.004772                                                         & 0.004815                                                             & 79.55\%            & 79.35\%                                                          \\ \hline
Con No-coadapt & 0.020719                                                          & 0.002584                                                         & 0.002583                                                             & 89.99\%            & 89.99\%                                                          \\ \hline
Con Coadapt    & 0.020719                                                          & 0.002583                                                         & 0.002405                                                             & 89.99\%            & 90.88\%                                                          \\ \hline
Pre-recov*      & 0.003834                                                          & 0.003105                                                         & 0.003077                                                             & 22.31\%            & 23.16\%                                                          \\ \hline
\end{tabular}
\caption{\label{tab:results}Losses and recovery. *Pre-recovery lesioned loss and percent recoveries based on the post-recovery values.}
\end{table}

@@

TODO: describe stopping conditions

\subsection{Experiment 1: AIP 75\% lesion}
This experiment demonstrates the ability for the co-processor to succeed after the
patient has lost some of the neural machinery responsible for representing the task.
In the non-coadaptive version of this experiment, the loss of AIP neurons involved
in encoding the visual inputs results in the simulated brain being unable to
condition its grasp for the object shape, though reaching largely still succeeds.
Likewise, the co-processor doesn't have sufficient information to improve performance.
In this case, the co-processor is effectively unable to improve on existing performance,
for lack of that information.

In the co-adaptive case, the brain and co-processor find a solution that largely recovers
performance. This demonstrates that the co-processor can adapt to the non-stationarity
of the mapping between the observed brain activity and the optimal stimulation. Note,
though, that the simulation is not designed to indicate if the co-processor speeds up
recovery, or results in greater results than natural recovery, since the simulation
is not designed to answer that question.

Additionally, in Fig. \ref{fig:results} we see that initially the co-processor and
brain did not strongly separate the object classes, but that they learned to as the
training proceeded. This is again due to the co-processor and brain learning how
to map the input visual information on to the task.

\subsection{Experiment 2: M1 75\% lesion}
TODO: after final results
* M1 + coadaptation shows we can adapt to changes in the output mappings
* Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
* Later: fine tuning and object class separation. Learning task structure.

\subsection{Experiment 3: F5 and M1 connectivity lesion}
As in the other cases, the co-processor's loss in this experiment improved
quickly, initially, but takes a long time to refine. During that refinement,
object differentiation occurs. Note that in this experiment, no information
can propagate forward to the M1 module, since we have completely severed the
input connections to that module. It is the sole responsibility of the
co-processor to perform that forward-propagation. As a result, co-adaptation
only affects the M1 module. The co-processor's learning algorithm successfully
adapted to this case, where the observed brain activity was stationary, but
the mapping between stimulation and task performance was not.

\subsection{Experiment 4: M1 lesion with pre-recovery}
In this case, we simulate some amount of recovery prior to the co-processor
